High-Dimensional Regularized Discriminant Analysis

### 2015-03-30 reports/paper-hdrda.Rnw
Applied some more edits to the paper.

results <- lapply(results, function(res) {
  res <- lapply(res, function(res_d) {
    errors <- do.call(cbind, res_d$errors)
    res_d[["errors"]] <- NULL
    cbind(do.call(cbind, res_d), errors)
  })
  do.call(rbind, res)
})
results <- data.frame(do.call(rbind, results), stringsAsFactors = FALSE)

results$lambda <- as.numeric(results$lambda)
results$gamma <- as.numeric(results$gamma)
results$Dudoit <- as.numeric(results$Dudoit)
results$HDRDA <- as.numeric(results$HDRDA)
results$Guo <- as.numeric(results$Guo)
results$Pang <- as.numeric(results$Pang)
results$Tong <- as.numeric(results$Tong)
results$Witten <- as.numeric(results$Witten)

results$d <- factor(results$d, labels = sort(unique(as.numeric(results$d))))

# Summary of the classification error rates for each data set and number of
# variables selected, d.
error_rates <- subset(results, select = -c(lambda, gamma))
m_error <- melt(error_rates, variable.name = "Classifier", value.name = "Error")
error_rates <- ddply(m_error, .(data_set, d, Classifier), plyr:::summarize,
                     Avg_Error = mean(Error), Num_Reps = length(Error))
error_rates <- ddply(error_rates, .(data_set, d, Classifier), transform,
                     SE_Error = sqrt(Avg_Error * (1 - Avg_Error) / Num_Reps))
error_rates <- subset(error_rates, select = -Num_Reps)
error_rates$Classifier <- factor(as.character(error_rates$Classifier))

max_se <- max(error_rates[["SE_Error"]])
@

@@

As a result, researchers have emphasized three standard approaches to modifying
the \emph{LDA} and \emph{QDA} classifiers for application to high-dimensional
data: 1) covariance-matrix regularization, 2) dimension reduction, and 3) model
parsimony. Covariance-matrix regularization methods are often based on the
well-known ridge-regression approach of
\cite{Hoerl:1970cd}. \cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd},
\cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud},
\cite{Ye:2006vx}, and \cite{Ye:2005uu} have proposed regularization methods to
improve the estimation of the class covariance matrices by adjusting the
eigenvalues of the sample covariance matrices to ensure positive
definiteness. Alternatively, dimension reduction typically transforms the
original data in the high-dimensional space to a subspace having a much smaller
dimension such that redundant and noisy information is removed
\citep{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Ye:2005hd,Antoniadis:2003bk}. Finally,
\cite{Mai:2012bf}, \cite{Fan:2012iq}, \cite{Merchante:2012vk},
\cite{Clemmensen:2011kr}, \cite{Huang:2010ju}, and \cite{Dudoit:2002ev} have
proposed restrictions on the population covariance matrices to reduce the number
of parameters that must be estimated. For instance, \cite{Dudoit:2002ev} have
assumed the features within each class are independent so that the population
covariance matrices are diagonal, resulting in positive-definite sample
covariance matrices. For excellent overviews of the regularization problem in
discriminant analysis and other techniques, see \cite{Murphy:2012uq} and
\cite{Mkhadri:1997gy}.

@@

We reparameterize the \emph{RDA} classifier similar to \cite{Hastie:2008dt} and
\cite{Halbe:2007ea} and refer to our parameterization as the high-dimensional
\emph{RDA} (\emph{HDRDA}) classifier. First, we show the pooling parameter in
the \emph{HDRDA} classifier determines the contribution of each training
observation to the estimation of each class covariance matrix, enabling
interpretability that has been previously lacking with the \emph{RDA} classifier
\citep{Bensmail:1996fw}. Next, applying a reasoning similar to \cite{Ye:2006tq},
we establish that the matrix operations corresponding to the null space of the
pooled sample covariance matrix are redundant and can be discarded from the
\emph{HDRDA} decision rule without loss of classificatory information. Moreover,
because the \emph{HDRDA} decision rule is invariant to adjustments to the zero
eigenvalues, the decision rule in the original feature space is equivalent to a
decision rule in a much lower dimension (i.e., the intrinsic dimension), such
that the matrix inverses and determinants of relatively small matrices can be
rapidly computed. Thus, the dimension reduction inherent in the \emph{HDRDA}
classifier yields a substantial reduction in computation during model
selection. Also, as a result, we show that a large number of shrinkage methods
that are special cases of the \emph{HDRDA} classifier have no effect on the zero
eigenvalues of the covariance-matrix estimators. Such methods include work from
\cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
\cite{Ramey:2013ji} and \cite{Xu:2009fl}.

@@

Several regularization methods, such as the methods considered by
\citet*{Xu:2009fl}, \citet*{Guo:2007te}, and \cite{Mkhadri:1995jp} have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and has reduced variability in the standard
bias-variance tradeoff. A common form of the covariance-matrix regularization
applies a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma) = \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm \Sigma}_k$ toward $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

Here, we define the \emph{HDRDA} classifier by first formulating the
covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
its clear interpretation as a linear combination of the crossproducts of the
training observations centered by their respective class sample means. We define
the convex combination
\begin{align}
  \widehat{\bm \Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
  \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
\end{align}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $ \bm x_i$ $(i =
1, \ldots, N)$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
\begin{align}
	\widehat{\bm \Sigma}_k(\lambda)
  &= \left( 1 - \lambda + \frac{\lambda n_k}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}}^K n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N
  I(y_i = k) \bm x_i \bm x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k)
  \bm x_i \bm x_i^{T} \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$. We
can interpret \eqref{eq:sig-lambda-alternative2} as a covariance-matrix
estimator that borrows from $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to
better estimate $ \bm \Sigma_k$. From \eqref{eq:sig-lambda-alternative2},
$\lambda$ weights the contribution of each of the $N$ observations in estimating
$\bm \Sigma_k$. Hence, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with
$N$ rather than $n_k$ observations.\footnote{TODO: Add more to the
  interpretation here.}

Below, we will show that the pooling operation is advantageous in increasing the
rank of each $\widehat{\bm \Sigma}_k(\lambda)$ from rank$(\widehat{\bm
  \Sigma}_k)$ to rank$(\widehat{\bm \Sigma})$ for $0 < \lambda \le 1$. Notice
that if $\lambda = 0$, then the observations from the remaining $K - 1$ classes
do not contribute to the estimation of $\bm \Sigma_k$, corresponding to
$\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights
$c_{ik}(\lambda)$ in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$,
corresponding to $\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we
define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm
  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly,
for $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
  \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k = \bm
X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the contours of
five multivariate normal populations for $\lambda = 0$ with unequal covariance
matrices. As $\lambda$ approaches 1, the contours become more similar, resulting
in identical contours for $\lambda = 1$.

\begin{figure}
<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', fig.show='hold'>>=
rda_contours(lambda = 0)
rda_contours(lambda = 1/4)
rda_contours(lambda = 3/4)
rda_contours(lambda = 1)
@ 
\caption{Contours of five multivariate normal populations as a function of the
  pooling parameter $\lambda$.}
\label{fig:hdrda-contours}
\end{figure}

As we have discussed above, several eigenvalue adjustment methods have been
proposed to increase eigenvalues (approximately) equal to 0. To further improve
the estimation of $\bm \Sigma_k$ and to stabilize the estimator's inverse, we
define the eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount
borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm \Sigma_k$, and the
\emph{shrinkage} parameter $\gamma$ determines the degree of eigenvalue
shrinkage. The choice of $\alpha_k$ allows for a flexible formulation of
covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to the
\emph{RDA} classifier from \cite{Friedman:1989tm}. Now, substituting
\eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the \emph{HDRDA} classifier
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm
  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
from a grid of candidate models via cross-validation \citep{Hastie:2008dt}. We
provide an implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the
{\tt sparsediscrim} R package, available on CRAN.

\subsection{Relationship with Other Classifiers}

The addition of the parameter $\alpha_k$ in \eqref{eq:hdrda-cov} is one of
convenience and allows the flexibility of various covariance-matrix estimators
proposed in the literature. In practice, we generally are not interested in
estimating $\alpha_k$ as it would incorporate $K$ additional tuning parameters
to estimate via cross-validation, which is counterproductive to our goal of
improving the \emph{RDA} classifier's efficiency. By setting $\alpha_k$ to a
predefined constant, the \emph{HDRDA} covariance-matrix estimator includes or
resembles a large family of estimators, some of which we mention below.

\subsubsection{\cite{Friedman:1989tm}}

\cite{Friedman:1989tm} proposed regularized discriminant analysis (RDA) by
incorporating a weighted average of the sample covariance matrix $\widehat{ \bm
  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm \Sigma}$ to
estimate $\bm \Sigma_k$ with
\begin{align}
    \widehat{\bm \Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad
    (k =
      1, \ldots, K),\label{eq:sig-lambda}
      \end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
\bm S_k = n_k \widehat{\bm \Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $ \bm
S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. If $\lambda = 0$,
\eqref{eq:sig-lambda} is equal to $\widehat{\bm \Sigma}_k$ used in
\eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces
to \eqref{eq:pooled-cov}.

To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
covariance-matrix estimator
\begin{align}
          \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm
            \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm
              \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
          \end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
established by substituting \eqref{eq:sig-rda} into
\eqref{eq:qda}.\footnote{TODO: Mention the values of $\alpha_k$ and $\gamma$
  needed to resemble HDRDA.}

\subsubsection{Ridge-Like Estimators}

TODO: List a few estimators here along with $\alpha_k = 1$ such that HDRDA
includes them all. For example, set $\gamma = \text{trace}(\widehat{\bm
  \Sigma}_k) / \min(n_k, p)$ for Srivastava and Kubokawa.

@@

\begin{lemma}\label{lemma:RDA-constant-term}
Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times
  1}$, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x - \xbar_{k'})$ $(1 \le k,
k' \le K)$.
\end{lemma}
\begin{proof}
Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that
$\bm U_2^{T} \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma
  1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{ \bm
  \Sigma})$ $(i = 1, \ldots, N)$, $\bm U_2^{T} \bm x_i = \bm 0_{(p-q) \times
  1}$.  Hence, $\bm 0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i \bm U_2^{T}\bm x_i
= \bm U_2^{T}(\xbar_k - \xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{
I(y_i = k) n_{k'} - I(y_i = k') n_k \}$. Therefore, $\bm U_2^{T} (\bm x -
\xbar_k) = \bm U_2^{T} (\bm x - \xbar_{k'})$.
\end{proof}

@@

Thus far, we have shown that our proposed classifier in \eqref{eq:hdrda} is
invariant to the term $\bm U_2$, thus yielding an equivalent classifier in
\eqref{eq:hdrda-decomposed} with a substantial reduction in computational
complexity. Here, we demonstrate that the computational efficiency in
calculating the inverse and determinant of $\bm W_k$ can be further improved via
standard matrix operations.

>>>>

max_se <- 1234567890
@

@@

As a result, researchers have emphasized three standard approaches to modifying
the \emph{LDA} and \emph{QDA} classifiers for application to high-dimensional
data: 1) covariance-matrix regularization, 2) dimension reduction, and 3)
parameter restrictions. Covariance-matrix regularization methods are often based
on the well-known ridge-regression approach of \cite{Hoerl:1970cd}.
\cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp},
\cite*{Guo:2007te}, \cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and
\cite{Ye:2005uu} have proposed regularization methods to improve the estimation
of the class covariance matrices by adjusting the eigenvalues of the sample
covariance matrices to ensure positive definiteness. Alternatively, dimension
reduction typically transforms the original data in the high-dimensional space
to a subspace having a much smaller dimension such that redundant and noisy
information is removed
\citep{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Ye:2005hd,Antoniadis:2003bk}. Finally,
\cite{Mai:2012bf}, \cite{Fan:2012iq}, \cite{Merchante:2012vk},
\cite{Clemmensen:2011kr}, \cite{Huang:2010ju}, and \cite{Dudoit:2002ev} have
proposed restrictions on the population covariance matrices to reduce the number
of parameters that must be estimated. For instance, \cite{Dudoit:2002ev} have
assumed the features within each class are independent so that the population
covariance matrices are diagonal, resulting in positive-definite sample
covariance matrices. For excellent overviews of the regularization problem in
discriminant analysis and other techniques, see \cite{Murphy:2012uq} and
\cite{Mkhadri:1997gy}.

@@

We reparameterize the \emph{RDA} classifier similar to \cite{Hastie:2008dt} and
\cite{Halbe:2007ea} and refer to our parameterization as the high-dimensional
\emph{RDA} (\emph{HDRDA}) classifier. First, we show the pooling parameter in
the \emph{HDRDA} classifier determines the contribution of each training
observation to the estimation of each class covariance matrix, enabling
interpretability that has been previously lacking with the \emph{RDA} classifier
\citep{Bensmail:1996fw}. Next, applying a reasoning similar to \cite{Ye:2006tq},
we establish that the matrix operations corresponding to the null space of the
pooled sample covariance matrix are redundant and can be discarded from the
\emph{HDRDA} decision rule without loss of classificatory information. Moreover,
because the \emph{HDRDA} decision rule is invariant to adjustments to the zero
eigenvalues, the decision rule in the original feature space is equivalent to a
decision rule in a much lower dimension (i.e., the intrinsic dimension), such
that the matrix inverses and determinants of relatively small matrices can be
rapidly computed. Thus, the dimension reduction inherent in the \emph{HDRDA}
classifier yields a substantial reduction in computation during model
selection. Also, as a result, we show that a large number of shrinkage methods
that are special cases of the \emph{HDRDA} classifier have no effect on the zero
eigenvalues of the covariance-matrix estimators. Such methods include work from
\cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
\cite{Ramey:2013ji} and \citet*{Xu:2009fl}.

@@

Several regularization methods, such as the methods considered by
\cite{Xu:2009fl}, \citet*{Guo:2007te}, and \cite{Mkhadri:1995jp} have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and has reduced variability in the standard
bias-variance tradeoff. A common form of the covariance-matrix regularization
applies a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma) = \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm \Sigma}_k$ toward $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

Here, we define the \emph{HDRDA} classifier by first formulating the
covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
its clear interpretation as a linear combination of the crossproducts of the
training observations centered by their respective class sample means. We define
the convex combination
\begin{align}
  \widehat{\bm \Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
  \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
\end{align}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $ \bm x_i$ $(i =
1, \ldots, N)$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
\begin{align}
	\widehat{\bm \Sigma}_k(\lambda)
  &= \left( 1 - \lambda + \frac{\lambda n_k}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}}^K n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N
  I(y_i = k) \bm x_i \bm x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k)
  \bm x_i \bm x_i^{T} \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
each of the $N$ observations in estimating $\bm \Sigma_k$ from all $K$ classes
rather than $n_k$ observations from a single class. As a result, we can
interpret \eqref{eq:sig-lambda-alternative2} as a covariance-matrix estimator
that borrows from $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to estimate $
\bm \Sigma_k$.

Below, we will show that the pooling operation is advantageous in increasing the
rank of each $\widehat{\bm \Sigma}_k(\lambda)$ from rank$(\widehat{\bm
  \Sigma}_k)$ to rank$(\widehat{\bm \Sigma})$ for $0 < \lambda \le 1$. Notice
that if $\lambda = 0$, then the observations from the remaining $K - 1$ classes
do not contribute to the estimation of $\bm \Sigma_k$, corresponding to
$\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights
$c_{ik}(\lambda)$ in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$,
corresponding to $\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we
define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm
  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma} = N^{-1} \bm X^{T} \bm X$. Similarly,
for $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
  \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k = n_k^{-1} \bm
X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the contours of
five multivariate normal populations for $\lambda = 0$ with unequal covariance
matrices. As $\lambda$ approaches 1, the contours become more similar, resulting
in identical contours for $\lambda = 1$.

\begin{figure}
<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', fig.show='hold'>>=
rda_contours(lambda = 0)
rda_contours(lambda = 0.2)
rda_contours(lambda = 0.4)
rda_contours(lambda = 0.6)
rda_contours(lambda = 0.8)
rda_contours(lambda = 1)
@ 
\caption{Contours of five multivariate normal populations as a function of the
  pooling parameter $\lambda$.}
\label{fig:hdrda-contours}
\end{figure}

As we have discussed above, several eigenvalue adjustment methods have been
proposed to increase eigenvalues (approximately) equal to 0. To further improve
the estimation of $\bm \Sigma_k$ and to stabilize the estimator's inverse, we
define the eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount
borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm \Sigma_k$, and the
\emph{shrinkage} parameter $\gamma$ determines the degree of eigenvalue
shrinkage. The choice of $\alpha_k$ allows for a flexible formulation of
covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to the
\emph{RDA} classifier from \cite{Friedman:1989tm}. Substituting
\eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the \emph{HDRDA} classifier
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm
  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
from a grid of candidate models via cross-validation. We provide an
implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the {\tt
  sparsediscrim} R package, available on CRAN.

The choice of $\alpha_k$ in \eqref{eq:hdrda-cov} is one of convenience and
allows the flexibility of various covariance-matrix estimators proposed in the
literature. In practice, we generally are not interested in estimating
$\alpha_k$ as it would incorporate $K$ additional tuning parameters to estimate
via cross-validation, which is counterproductive to our goal of improving the
\emph{RDA} classifier's efficiency. By setting $\alpha_k$ to an appropriate
constant, the \emph{HDRDA} covariance-matrix estimator includes or resembles a
large family of estimators. Notice that if $\alpha_k = 1$ and $\lambda = 1$,
\eqref{eq:hdrda-cov} is equivalent to the standard ridge-like covariance-matrix
estimator in \eqref{eq:ridge-estimator}, and various estimators proposed in the
literature can be obtained by selecting $\gamma$ accordingly. For instance, with
$\gamma = \tr\{ \widehat{\bm \Sigma} \} / \min(N, p)$, we obtain the estimator
from \cite{Srivastava:2007ww}.

Perhaps the most well-known of the ridge-like estimators is from
\cite{Friedman:1989tm}, who proposed regularized discriminant analysis (RDA) by
incorporating a weighted average of the sample covariance matrix $\widehat{ \bm
  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm \Sigma}$ to
estimate $\bm \Sigma_k$ with
\begin{align}
    \widehat{\bm \Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad
    (k =
      1, \ldots, K),\label{eq:sig-lambda}
      \end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
\bm S_k = n_k \widehat{\bm \Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $ \bm
S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. If $\lambda = 0$,
\eqref{eq:sig-lambda} is equal to $\widehat{\bm \Sigma}_k$ used in
\eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces
to \eqref{eq:pooled-cov}. To further improve the estimation of $\bm \Sigma_k$
and to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm}
has proposed the biased covariance-matrix estimator
\begin{align}
          \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm
            \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm
              \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
          \end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
established by substituting \eqref{eq:sig-rda} into \eqref{eq:qda}.

Our proposed \emph{HDRDA} classifier resembles \eqref{eq:sig-rda} but improves
the interpretation of the pooling parameter $\lambda$ as discussed
above. Moreover, as we discuss below, our formulation of \emph{HDRDA} classifier
lends itself to rapid estimation in high-dimensional classification.

@@

\begin{lemma}\label{lemma:RDA-constant-term}
Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times
  1}$, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x - \xbar_{k'})$ $(1 \le k,
k' \le K)$.
\end{lemma}
\begin{proof}
Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that
$\bm U_2^{T} \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma
  1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{ \bm
  \Sigma})$ $(i = 1, \ldots, N)$, $\bm U_2^{T} \bm x_i = \bm 0_{p-q}$.  Hence,
$\bm 0_{p-q} = \sum_{i=1}^N \beta_i \bm U_2^{T}\bm x_i = \bm U_2^{T}(\xbar_k -
\xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i =
k') n_k \}$. Therefore, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x -
\xbar_{k'})$.
\end{proof}

@@

As we can see, we are able to avoid the time-consuming inverses and determinants
of $p \times p$ covariance matrices in \eqref{eq:hdrda} and instead calculating
these same operations on $\bm W_k \in \mathbb{R}_{q \times q}$ in
\eqref{eq:hdrda-decomposed}. The substantial computational improvements arise
because our proposed classifier in \eqref{eq:hdrda} is invariant to the term
$\bm U_2$, thus yielding an equivalent classifier in \eqref{eq:hdrda-decomposed}
with a substantial reduction in computational complexity. Here, we demonstrate
that the computational efficiency in calculating the inverse and determinant of
$\bm W_k$ can be further improved via standard matrix operations by showing that
the inverses and determinants of $\bm W_k$ can be performed on matrices of size
$n_k \times n_k$.
